---
title: "COVID-19 in New York State"
subtitle: "Source: https://coronavirus.health.ny.gov/county-county-breakdown-positive-cases"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    social: menu
    theme: readable
    vertical_layout: fill
date: 'Last Updated: `r Sys.time()` UTC'
runtime: shiny
resource_files:
- data/covid_count_map/covid_count.dbf
- data/covid_count_map/covid_count.prj
- data/covid_count_map/covid_count.shx
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
# Load libraries needed

library(flexdashboard)
library(knitr)
library(shiny)
library(mapview)
library(leaflet)
library(dichromat)
library(ggthemr)
library(plotly)
library(lubridate)
library(DT)
library(data.table)
library(sp)
library(sf)
library(tidyverse)
library(janitor)

# Load datasets
covid_map <- st_read("data/covid_count_map/covid_count.shp") %>%
  rename(
    "positive_cases" = "pstv_cs",
    "name_nyc" = "nam_nyc"
  )

covid_time_series <- fread("data/covid_time_series/covid_time_series.csv") %>%
  clean_names() %>%
  mutate(date = ymd(date))

top_7 <- covid_time_series %>%
  filter(date == max(date)) %>%
  arrange(desc(positive_cases)) %>%
  head(7) %>%
  select(county) %>%
  pull(county)

covid_time_series <- covid_time_series %>%
  mutate(county = ifelse(county %in% top_7, as.character(county), "Other")) %>%
  group_by(county, date) %>%
  summarise(positive_cases = sum(positive_cases)) %>%
  ungroup()

levels <- covid_time_series %>%
  filter(date == max(date)) %>%
  arrange(desc(positive_cases)) %>%
  select(county) %>%
  pull(county)

covid_time_series$county <- factor(covid_time_series$county, levels = levels)
```


Dashboard
====================================================================== 

Data source: [New York State Department of Health](https://coronavirus.health.ny.gov/county-county-breakdown-positive-cases)

Column {data-width=650, data-height = 200}
-----------------------------------------------------------------------

###

```{r}
# Include theme for plots
ggthemr("dust")

county_cases_counties <- covid_time_series %>%
  ggplot(aes(x = date, y = positive_cases, group = county, colour = county)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Date",
    y = "Positive Cases",
    title = "Cumulative COVID-19 in New York State counties"
  )

ggplotly(county_cases_counties)
```

### 

```{r}
ggthemr("dust")

covid_ny <- covid_time_series %>%
  select(date, positive_cases) %>%
  group_by(date) %>%
  summarise(positive_cases = sum(positive_cases)) %>%
  mutate(new_cases = positive_cases - lag(positive_cases)) %>%
  ggplot(aes(x = date, y = positive_cases)) +
  geom_line() +
  geom_point() +
  geom_bar(aes(y = new_cases), stat = "identity") +
  labs(
    x = "Date",
    y = "Positive Cases",
    title = "Cumulative COVID-19 in New York State"
  ) +
  theme(legend.position = "right")

ggplotly(covid_ny)
```


Column {data-width = 450}
-----------------------------------------------------------------------

### Covid-19 Confirmed Cases Map 

```{r,message = FALSE, message = FALSE, warning = FALSE}

pal <- colorRampPalette(c("#fae6c0", "#7A6752"))
m <- mapview(covid_map,
  zcol = "positive_cases",
  col.regions = pal,
  alpha = 0.8,
  alpha.regions = 0.9
)

m@map %>% setView(lng = -74, lat = 42, zoom = 6.5)
```


Data
====================================================================== 

Column {data-width = 450}
-----------------------------------------------------------------------

###

```{r}

# Time series of covid-19 cases in NY State
covid_series_dataset <- fread("data/covid_time_series/covid_time_series.csv") %>%
  clean_names() %>%
  mutate(date = ymd(date))
# Define UI

ui <- fluidPage(
  titlePanel("COVID-19 positive cases in New York State"),

  # Rows for selected inputs
  fluidRow(
    column(
      2,
      selectInput(
        "county",
        "County:",
        c("All", unique(as.character(covid_series_dataset$county)))
      )
    ),
    column(
      2,
      selectInput(
        "date",
        "Date:",
        c("All", unique(as.character(covid_series_dataset$date)))
      )
    ),
    # Create a new row for the table
    dataTableOutput("table")
  )
)

# Define server logic
server <-
  function(input, output) {

    # Filter data based on selected items
    output$table <- renderDataTable(DT::datatable({
      data <- covid_series_dataset
      if (input$county != "All") {
        data <- data[data$county == input$county, ]
      }
      if (input$date != "All") {
        data <- data[as.Date(data$date) == input$date, ]
      }
      data
    }),  options = list(pageLength = 25))
  }

# Run the app
shinyApp(ui = ui, server = server)
```

Column {data-width = 450}
-----------------------------------------------------------------------

### 

Hi, 

My name is Sue Marquez. I am a Data Scientist based in New York City. At the beggining of COVID-19 epidemic in the US I started collecting daily updates on the number of cases testing positive for the COVID-19 in the state of New York. 

On a daily basis the [NY Department of Health](https://coronavirus.health.ny.gov/county-county-breakdown-positive-cases) was updating the cumulative count of cases testing positive for COVID-19 at the county level. During the crisis the NYS COVID website WAS NOT publicly reporting:

1. Cases at the borough level for New York City. 
2. Daily count of number of COVID tests performed.

Given the cumulative count, it was hard to know the change in number cases from day to day, reason why I started scraping the website on a daily basis. 

If you are interested in obtaining the csv file, you can either download it below or access my [github repository](https://github.com/Suemarquez/covid_nys), which also includes associated code. 




```{r}
# Dowload data 
ui <- fluidPage(
  titlePanel("Download Data"),
  
  # Button
  downloadButton("downloadData", "Download")
)

server <- function(input, output) {
  
  datasetInput <- fread("data/covid_time_series/covid_time_series.csv")
  
  output$downloadData <- downloadHandler(
    filename = "covid_nys_time_series.csv",
    content = function(file) {
      write.csv(datasetInput, file, row.names = FALSE)
    }
  )
}

# Run the app
shinyApp(ui = ui, server = server)
```


